Wi-Fi Based Handwritten Signature Verification Using a Triplet Network

### 2019-09-03 190410_Paper_ACPR/paper_v7_r5-1.tex
ACPR_paper: rev 7

\begin{document}
%
\title{Deep Triplet Network adopting the Kernel and the Range Space Learning for Wi-Fi Handwritten Signature Verification}
%
\titlerunning{Triplet Network adopting KAR Learning for Wi-Fi Signature Verification}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Young-Woong Kwon \and Jooyoung Kim \and Kar-Ann Toh}
%
\authorrunning{Y. Kwon et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
School of Electrical and Electronic Engineering, Yonsei University, 50 Yonsei-ro, Seodaemun-gu, Seoul 03722, Republic of Korea\\
\email{\{herokwon, harrykim, katoh\}@yonsei.ac.kr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
In this paper, we propose a system for identity verification based on the handwritten signature signals captured by the Wi-Fi CSI signals. 
The kernel and the range space learning is initially adopted for refining the triplet inputs for fast loss convergence by mining the distinctive inputs from the training Wi-Fi signature signals. 
Subsequently, the triplet network utilizing the ConvNet structure is trained with the mined triplet inputs based on L-2 distance comparison. 
Our experiments on an in-house Wi-Fi handwritten signature signal dataset show encouraging verification accuracy with faster training loss convergence comparing with the baseline triplet network and the Siamese network.

@@

In recent years, several behavioral biometric traits for identity authentication have been investigated in view of their rigid physical body independence~\cite{bailador2011analysis}. Among the behavioral biometrics, the signature-based user authentication~\cite{sanmorino2012survey,galbally2015line} is attracting considerable interest with the development of in-air signature recognition systems~\cite{jeon2012system,sesa2012information,malik20183dairsig}.
With the help of sensors such as the depth camera~\cite{malik20183dairsig} or mobile sensor~\cite{jeon2012system}, the in-air signature recognition system has lower spatial constraint in the signature acquisition process than the contact-based authentication systems.

Recently, the commercial Wi-Fi device has been adopted for in-air signature authentication due to its easy accessible property \cite{moon2017air}. Based on the distortion of the  Wi-Fi CSI signal according to the user's gestures, the in-air signature recognition system showed reasonable user verification performance~\cite{moon2017air}.

More recently, some studies attempted to implement the deep learning algorithms in Wi-Fi signal-based user authentication systems to improve the verification performance ~\cite{shi2017smart,pokkunuru2018neuralwave}.

In this paper, we utilize the deep triplet network for identity verification based on the Wi-Fi CSI signature signal. To achieve not only the desired verification accuracy but also a fast training speed, we adopt the kernel and the range (KAR) space leaning~\cite{toh100,toh2018learning,toh2018analytic,toh2018gradient} in order to mine the distinctive triplet inputs. Subsequently, the triplet network utilizing the ConvNet structure as a feature extractor is trained based on the L-2 distance comparison.

The main contributions of our work can be summarized as follows:
\begin{itemize}
\item Proposal of a system for identity verification based on the Wi-Fi handwritten signature signals using a deep triplet network.
\item Adopted the kernel and the range (KAR) space learning in order to mine the distinctive triplet inputs which boost the convergence speed of the training loss in triplet network.
\item Provision of the experimental study on an in-house Wi-Fi handwritten signature signal dataset collected from 50 subjects.
\end{itemize}

The paper is organized as follows: related works including the triplet network and KAR space learning will be introduced in Section 2 for immediate reference. Our proposed method will be discussed in Section 3. Section 4 describes our experimental results and analysis. Some concluding remarks will be given in Section 5.

@@

\subsection{Triplet network}
% triplet network
The triplet network is considered a metric learning based model~\cite{weinberger2006distance} which aims to learn useful representations by means of distance comparison~\cite{hoffer2015deep}. 
\textbf{It is often seen in person re-identification~\cite{chen2017beyond,cheng2016person,ding2015deep,schroff2015facenet,wang2016joint} where the individual identities are matched based on discriminative image features. The main difference between person re-identification and identification is the person re-identification is more challenging task by associating images of the same person taken from different cameras or from the same camera in different occasions. }

In these problem, to distinguish among the classes is a challenging task since the features are relatively weak compared to the background features. To address this problem, the triplet network optimizes the embedding data space such that those data points with the same identity are closer to each other than those with different identities~\cite{hermans2017defense}.

The triplet network receives triplet pairs as its input. These triplets pairs are constructed based on a combination of the input data. Since not all triplet samples contribute to the training, recent attention~\cite{schroff2015facenet} has been paid to the choice of relevant input pairs for training. To optimize the training process which utilizes only some parts of the triplet pairs, several researches~\cite{cheng2016person,ding2015deep,wang2016joint} generated triplet \textbf{from a small number of classes (persons) in each iteration.}

Recently, \cite{schroff2015facenet} adopted a triplet mining process to speed up the training convergence. They utilized large mini-batch at each training iteration and selected the triplet based on the training network rather than the random sampling. However, this strategy needed a few thousands of exemplars mini-batch in every training iteration to select the triplet pairs. This results in a heavy computational load in training.

% Rel works: KAR learning
\subsection{Kernel and the range space learning}\label{kar}

The multilayer feedforward neural networks is generally trained by the gradient descent method based on backpropagation \cite{goodfellow2016deep}.
However, setting the learning parameters such as learning rate or momentum value is time consuming and yet important to arrive at reasonable solution.

@@

% Karnet structure and mining samples.
Let $\mathbf{X}\in{\mathbb{R}}^{m \times (n+1)}$ denotes the training data set and $\mathbf{G}\in{\mathbb{R}}^{m \times n}$ denotes the network outputs.
Then the multilayer neural network structure can be written in linear equation form as follows:
\begin{equation}
    \mathbf{G} = \sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right), i=1,\dotsc,n,
\end{equation}
where $\mathbf{W}_{1}\in{\mathbb{R}}^{(n+1) \times h_{1}}$,$\mathbf{W}_{2}\in{\mathbb{R}}^{(h_{1}+1) \times h_{2}}$,$\dots,\mathbf{W}_{i}\in{\mathbb{R}}^{(h_{(i-1)}+1) \times n}$,$\mathbf{1}=\left[1,\dots,1\right]^{T}\\
\in{\mathbb{R}}^{m \times 1}$ and $\sigma(.)$ is activation function.
By adopting the one-hot encoded target $\mathbf{Y}\in{\mathbb{R}}^{m \times n}$ instead of network output $\mathbf{G}$, The trained weight matrices $\mathbf{W}_{i}$ using KAR space manipulation can be obtained as follows~\cite{toh2018gradient}:

@@

% Methods: System overview(Fig.1)
In this section, we propose an identity verification system based on the Wi-Fi in-air handwritten signature (which will be called Wi-Fi signature signals hereafter) using the triplet network~\cite{hoffer2015deep}. Fig.1 shows an overview of the proposed system utilizing the kernel and the range (KAR) space learning~\cite{toh2018learning,toh2018gradient} for the triplet input mining together with the triplet network.
Essentially, the KAR space projection learning is trained and utilized to generate the triplet input data by mining the hard positive and the hard negative samples from each given anchor sample in the training dataset (see item (a) in Fig.~\ref{fig1}). \textbf{The hard positive and hard negative samples refer to positive and negative class data which are likely to be misclassified by the network.}
Subsequently, the ConvNet structure in the triplet network (see item (b) in Fig.~\ref{fig1}) is trained with the mined triplet data based on the triplet loss function using L-2 distance comparison (see item (c) in Fig.~\ref{fig1}).
The following subsections describe the details of the triplet mining using KAR space learning and the triplet network.

@@

The triplet network receives a triplet of data as its inputs. These triplet data set consists of the reference data (will be called anchor samples hereafter) and the corresponding positive class (same class with that of the anchor) and negative class data (different class from that of the anchor). \textbf{The goal of the triplet network is to position the feature vectors in appropriate separation space by putting the positive sample closer to the anchor sample and keeping the negative sample away from the anchor sample.}

\textbf{According to \cite{schroff2015facenet}, it is important to select the hard positive and the hard negative samples from each given anchor sample for the faster loss convergence when training the triplet network.
The hard positive sample denotes the positive sample whose distance to the anchor sample is greater than others (which is most likely to be misclassified as a negative sample) while the hard negative sample denotes the negative sample whose distance to the anchor sample is smaller (which is most likely to be misclassified as a positive sample). However, there is no information about which sample is the hard positive or negative before we train the network.}

@@

By training the network with using the single shot KAR space learning method, we can calculate the L-2 distance between every training samples by using the output vector from the KAR space network. Taking each training sample as an anchor sample $\mathbf{x}_{anc}$, to measure the distance between $\mathbf{x}_{anc}$ and other samples, the trained network output given by:

\begin{equation}
    f\left(\mathbf{X}\right) = \sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\cdot\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right),
\end{equation}

is used to generate the hard positive and hard negative samples.
The hard positive sample $\mathbf{x}_{pos}$ and the hard negative sample $\mathbf{x}_{neg}$ with respect to the anchor sample $\mathbf{x}_{anc}$ can be respectively to determined by:

\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{pos}\right)}} \right\|_2^2} \geq \mathrm{t}_{pos}, \label{pos}
\end{equation}
\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{neg}\right)}} \right\|_2^2} \leq \mathrm{t}_{neg},\label{neg}
\end{equation}

where $\mathrm{t}_{pos}$ and $\mathrm{t}_{neg}$ denotes the thresholds which determines whether a sample is hard positive/negative. Here we utilized equations~\eqref{pos} and~\eqref{neg} to select the hard positive/negative samples instead of just using the hardest positive/negative samples since the hardest samples are likely to be outliers which degrade the training process of the triplet network.
We empirically set the $\mathrm{t}_{pos}$ to 75 percentile of L-2 distance and $\mathrm{t}_{neg}$ to 25 percentile of L-2 distance and the final $\mathbf{x}_{pos}$ and $\mathbf{x}_{neg}$ are randomly selected among the samples which match the equations.

% Methods: ConvNets
\subsection{ConvNet structures}

To design the triplet network, we first need to select the feature extractor which converts the input triplet data into feature vectors. In this work, we utilize the ConvNet structure~\cite{lecun1998gradient} as a feature extractor since the three-dimensional data format of our preprocessed input signal can be regarded as an image data format with multiple channels.

Our ConvNet structure (item (b) in Fig~\ref{fig1}) for the network consists of $i$ convolutional layers $\mathbf{C}_{i}$ and one fully-connected layer $\mathbf{F}$. The number of convolutional filters to be trained in each layer is empirically chosen as $\{64, 128, ...,  2^{6+i}\}$, with fixed filter size of $3\times3$ and stride of 1. The Rectiﬁed Linear (ReLU) function as an activation function and the Max-pooling layer are applied between each convolutional layer. Subsequently, the features from the last convolutional layer are directly flattened into a vector before the fully-connected layer $\mathbf{F}$.
The output vectors from the fully-connected layer are finally transformed using the sigmoid function following with the L-2 normalization.

%Since the networks utilize three ConvNet structures which ties the weights each other, noting here that three structures described in Fig~\ref{fig1} (b) are actually the same model.

% Methods: Triplet loss
\subsection{Triplet loss}

The triplet loss function was first seen in \cite{hoffer2015deep} to train the triplet network.
For the $i_{th}$ anchor input sample $\mathbf{x}_{anc,i}$, the triplet input is generated by grouping the hard positive input samples $\mathbf{x}_{pos,i}$ and the hard negative input samples $\mathbf{x}_{neg,i}$ selected based on the KAR space learning. The generated triplet input $\left\{\mathbf{x}_{anc,i},\mathbf{x}_{pos,i},\mathbf{x}_{neg,i}\right\}$ is then respectively transformed to the anchor, the positive and the negative feature vectors $\left\{\mathbf{v}_{anc,i},\mathbf{v}_{pos,i},\mathbf{v}_{neg,i}\right\}$ with the ConvNet structure.

Here, the triplet loss function is calculated by comparing the positive distance (L-2 distance between the anchor vector and the positive vector) and the negative distance (L-2 distance between the anchor vector and the negative vector) as follows:

\begin{equation}
    loss = \sum_i^N max\left({ \left[ {\left\| {{\mathbf{v}_{anc,i}} - {\mathbf{v}_{pos,i}}} \right\|_2^2} - {\left\| {{\mathbf{v}_{anc,i}} - {\mathbf{v}_{neg,i}}} \right\|_2^2}  + \alpha \right]}, 0 \right),\label{triplet}
\end{equation}

where $N$ denotes the size of the mini-batch, ${\left\| . \right\|_2^2}$ denotes the L-2 distance and $\alpha$ denotes the preset margin.
The ConvNet structure using equation~\eqref{triplet} is thus trained to maximize the gap between the positive distance and the negative distance which should be larger than the margin $\alpha$.

@@

% Dataset
\subsection{Dataset}
 In order to evaluate the verification performance of the proposed system, the Wi-Fi CSI signature dataset \cite{moon2017air} with single position was utilized in our experiments. After adopting the data preprocessing process proposed in \cite{moon2017air}, the Wi-Fi CSI signature dataset consists of 2000 Wi-Fi CSI signature signals (4 directions $\times$ 10 samples $\times$ 50 identities) with the sample size of 500$\times$30$\times$6. We utilized only the absolute value from each complex CSI signal in our experiments since the 2.4Ghz Wi-Fi CSI signals have issues of device firmware in their phase signal~\cite{wang2015understanding}.

@@

\subsubsection{Performance evaluation}
The proposed system is evaluated under two cases: i) case I on comparison between the proposed system and other linear or deep learning-based methods based on verification accuracy, and case II on detailed comparison between the proposed system and the deep learning-based methods using receiver operating characteristic (ROC) curve and training loss curve. For the Case I, existing linear methods such as the least squares error estimation (LSE), the principal components analysis (PCA) [22] with LSE, the support vector machine (SVM) with different kernel function and total error rate minimization which adopts the reduced multivariate polynomial model as basis function (TER-RM2) \cite{toh2003fingerprint,toh2008between}, the deep learning-based Siamese network \cite{koch2015siamese} and the baseline triplet network \cite{hoffer2015deep} are included for performance benchmarking. The Siamese network and the baseline triplet network utilized the same ConvNet structures with the proposed system and only differed in their input data style and loss function.

The verification performance of the proposed system and other methods were evaluated in terms of Equal Error Rates (EER, \%) averaged from five runs of two-fold cross-validation test. Due to the memory constraint caused by the large data size, noting here that the deep learning-based methods utilized randomly sampled 9,500 negative pairs in the validation stage, which is the same number as the number of the positive pairs.

\subsubsection{Network Structure and Parameter Settings}
The multilayer feedforward network structure of KAR space learning is specified in Table~\ref{tab2}. With input data, size of 500$\times$30$\times$6, we set two network layers where the size of the each layer is 1024 and 16, respectively. Each layer is initialized with uniform distribution over [0, 1). We used $\sigma = {tan}^{-1}$ as an activation function following \cite{toh2018analytic}.

\begin{table}[]
    \caption{The network structure of the KAR space learning.}\label{tab2}
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
    Layer   & Size     & Activation \\ \hline
    Input   & 500$\times$30$\times$6 &            \\
    Fully-Connected 1 & 1$\times$1$\times$1024 & $\sigma = {tan}^{-1}$     \\
    Fully-Connected 2 & 1$\times$1$\times$16  & $\sigma = {tan}^{-1}$     \\
    Output  & 1$\times$1$\times$50   &            \\ \hline
    \end{tabular}
\end{table}

For the proposed system and the deep learning-based methods, we utilized the same ConvNet structure as specified in Table~\ref{tab1}. We trained the network starting with a learning rate of 0.00005 and a mini-batch, size of 32. We optimized the loss by the Adam optimizer with L-2 penalty of 0.0002 except for the output layer. The output layer is regularized with L-2 penalty of 0.0001. We initialized all network weights in the convolutional layers with normal distribution of zero-mean and standard deviation of 0.01. Biases were also initialized with normal distribution of 0.5 mean and standard deviation of 0.01. For the triplet networks, the hyper-parameter regulating triplet loss is empirically set to 0.1. The training epochs are set to 1,500 for all three deep learning-based algorithms. For the linear methods such as LSE,SVM and TER, the input signal is initially resized to 500 $\times$ 30 by averaging along the subcarrier axes due to limitation of hardware memory. For the PCA-LSE, the input dimension is reduced to 40.

>>>>

\begin{document}
%
\title{Wi-Fi Based Handwritten Signature Verification Using a Triplet Network}
%
\titlerunning{Wi-Fi Based Handwritten Signature Verification Using a Triplet Network}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Young-Woong Kwon \and Jooyoung Kim \and Kar-Ann Toh}
%
\authorrunning{Y. Kwon et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
School of Electrical and Electronic Engineering, Yonsei University, 50 Yonsei-ro, Seodaemun-gu, Seoul 03722, Republic of Korea\\
\email{\{herokwon, harrykim, katoh\}@yonsei.ac.kr}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in 15--250 words.
In this paper, we propose a system for identity verification based on the handwritten signature signals captured by the Wi-Fi Channel State Information (CSI). 
To enable a fast loss convergence, a kernel and the range space learning is initially adopted for refining the triplet inputs by mining the distinctive inputs. 
Subsequently, the triplet network is trained on a ConvNet structure using the mined triplet inputs. 
Our experiments on an in-house Wi-Fi handwritten signature signal dataset show encouraging verification accuracy with faster training loss convergence comparing with the baseline triplet network and the Siamese network.

@@

Over recent years, several behavioral biometric traits attracted attention in view of their rigid physical body independence~\cite{bailador2011analysis}. Among these behavioral biometrics, the signature-based user authentication~\cite{sanmorino2012survey,galbally2015line} has attracted considerable interest with the development of in-air signature recognition systems~\cite{jeon2012system,sesa2012information,malik20183dairsig}.
With the help of sensors such as the depth camera~\cite{malik20183dairsig} or a mobile sensor~\cite{jeon2012system}, the in-air signature recognition system has lower the spatial constraint in the process of signature acquisition comparing with contact-based authentication systems.

Recently, the commercial Wi-Fi device has been adopted for in-air signature authentication due to its easy accessible property \cite{moon2017air}. Based on the distortion of the  Wi-Fi CSI signal according to the user's gestures, the in-air signature recognition system showed reasonable user verification performance~\cite{moon2017air}. More recently, some studies attempted to implement the deep learning algorithms in Wi-Fi signal-based user authentication systems to improve the verification performance ~\cite{shi2017smart,pokkunuru2018neuralwave}.

In this paper, we utilize the deep triplet network for identity verification based on the Wi-Fi CSI signature signal. To achieve not only the desired verification accuracy but also a fast training speed, we adopt the kernel and the range (KAR) space learning~\cite{toh100,toh2018learning,toh2018analytic,toh2018gradient} in order to mine the distinctive triplet inputs. Subsequently, the triplet network which utilizes the ConvNet structure as a feature extractor is trained based on the L-2 distance comparison.

The main contributions of our work can be summarized as follows:
\begin{itemize}
\item Proposal of a system for identity verification based on the Wi-Fi handwritten signature signals using a deep triplet network.
\item Adopted the kernel and the range (KAR) space learning in order to mine the distinctive triplet inputs which boost the convergence speed of the training loss in the triplet network.
\item Provision of an experimental study using an in-house Wi-Fi handwritten signature dataset collected from 50 subjects.
\end{itemize}

The paper is organized as follows: related works including the triplet network and KAR space learning are introduced in Section 2 for immediate reference. Our proposed method are discussed in Section 3. Section 4 describes our experimental results and analysis. Some concluding remarks are given in Section 5.

@@

\subsection{Triplet network}
% triplet network
The triplet network is considered a metric learning based model~\cite{weinberger2006distance} which aims to learn useful representations by means of distance comparison~\cite{hoffer2015deep}. 
It is often seen in person re-identification~\cite{chen2017beyond,cheng2016person,ding2015deep,schroff2015facenet,wang2016joint} where the individual identities are matched based on discriminative image features. The main difference between person identification and re-identification is that the later is a more challenging task where images of the same person taken from different cameras or from the same camera in different occasions are to be associated. In this re-identification problem, to distinguish among the classes is a challenging task since the features are relatively weak compared with the background features. To address this problem, we adopt a triplet network which optimizes the embedding data space so that data points with the same identity are closer to each other than those with different identities~\cite{hermans2017defense}.

The triplet network receives triplet pairs of data as its input. These data triplets are constructed based on a combination of the input data. Since not all triplet samples contribute to the desired classification, recent attention~\cite{schroff2015facenet} has been paid to the choice of relevant input pairs for training. In order to optimize the training process which utilizes only some parts of the triplet pairs, several researches~\cite{cheng2016person,ding2015deep,wang2016joint} generated triplet from a small number of classes (persons) in each iteration. Recently, \cite{schroff2015facenet} adopted a triplet mining process to speed up the training convergence. They utilized a large mini-batch at each training iteration and selected the triplet based on the training network rather than random sampling. However, this strategy needed a few thousands of exemplar mini-batches in every training iteration for triplet pairs selection. This results in a heavy computational load in training.

% Rel works: KAR learning
\subsection{Kernel and the range space learning~\cite{toh100,toh2018learning,toh2018analytic,toh2018gradient}}\label{kar}

Generally, the multilayer feedforward neural networks is trained based on the gradient descent method via backpropagation~\cite{goodfellow2016deep}.
However, setting the learning parameters such as the learning rate or the learning momentum is a time consuming task.

@@

% Karnet structure and mining samples.
Given $m$ training samples. Let $\mathbf{X}\in{\mathbb{R}}^{m \times (n+1)}$ denotes the training data set and $\mathbf{G}\in{\mathbb{R}}^{m \times n}$ denotes the network outputs.
Then the multilayer neural network structure can be written in linear equation form as follows:
\begin{equation}
    \mathbf{G} = \sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right), i=1,\dotsc,n,
\end{equation}
where $\mathbf{W}_{1}\in{\mathbb{R}}^{(n+1) \times h_{1}}$,$\mathbf{W}_{2}\in{\mathbb{R}}^{(h_{1}+1) \times h_{2}}$,$\dots,\mathbf{W}_{i}\in{\mathbb{R}}^{(h_{(i-1)}+1) \times n}$,$\mathbf{1}=\left[1,\dots,1\right]^{T}\\
\in{\mathbb{R}}^{m \times 1}$ and $\sigma(.)$ is the activation function.
By adopting an one-hot encoded target $\mathbf{Y}\in{\mathbb{R}}^{m \times n}$, training of the weight matrices $\mathbf{W}_{i}$ using the KAR space method~\cite{toh2018gradient} can be computed as follows:

@@

% Methods: System overview(Fig.1)
In this section, we propose an identity verification system based on the Wi-Fi in-air handwritten signature (which will be called Wi-Fi signature hereafter) using the triplet network~\cite{hoffer2015deep}. Fig.1 shows an overview of the proposed system utilizing the kernel and the range (KAR) space learning~\cite{toh2018learning,toh2018gradient} for mining the triplet input.
Essentially, the KAR space projection learning is utilized to learn the triplet input data by mining the hard positive and the hard negative samples from each given anchor sample (see item (a) in Fig.~\ref{fig1}). The hard positive and the hard negative samples refer to positive and negative class samples which are likely to be misclassified by the network.
Subsequently, the ConvNet structure in the triplet network (see item (b) in Fig.~\ref{fig1}) is trained with the mined triplet data based on a triplet loss function using the L-2 distance comparison (see item (c) in Fig.~\ref{fig1}).
The following subsections describe the details of the triplet mining using KAR space learning and the triplet network.

@@

The network receives a triplet set of data as its inputs. These triplet data consist of the reference data (will be called anchor samples hereafter) and the corresponding positive class data (same class with that of the anchor) and negative class data (different class from that of the anchor). The goal of the triplet network is to position the feature vectors with appropriate separation space by putting the positive samples close to the anchor sample and keeping the negative sample away from the anchor sample.

According to \cite{schroff2015facenet}, it is important to select the hard positive samples and the hard negative samples with reference to the given anchor sample for fast loss convergence when training the triplet network.
A hard positive sample is defined as a sample whose distance to the anchor sample is large (which is most likely to be misclassified as a negative sample). On the other hand, a hard negative sample is defined as a sample whose distance to the anchor sample is small (which is most likely to be misclassified as a positive sample). However, from the raw data, there is no information about regarding whether a sample is considered hard positive or hard negative before we train the network.

@@

By training the network with the single shot KAR space learning, we can map the L-2 distance between every samples by using the output vector of the KAR space network. \textbf{For training dataset $\mathbf{X}$, the trained network output is given by:}
\begin{equation}
    f\left(\mathbf{X}\right) = \sigma\left(\left[\mathbf{1},\sigma\left(\dots\left[\mathbf{1},\sigma\left(\left[\mathbf{1},\sigma\left(\mathbf{X}\cdot\mathbf{W}_{1}\right)\right]\mathbf{W}_{2}\right)\right]\dots\mathbf{W}_{(i-1)}\right)\right]\mathbf{W}_{i}\right),
\end{equation}
\textbf{For anchor sample $\mathbf{x}_{anc}$ which is randomly selected from training dataset X, the output of the KAR learning makes it possible to measure the distance between each sample.}
A hard positive sample $\mathbf{x}_{pos}$ and a hard negative sample $\mathbf{x}_{neg}$ with respect to the anchor sample $\mathbf{x}_{anc}$ can each be respectively determined based on:
\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{pos}\right)}} \right\|_2^2} \geq \mathrm{t}_{pos}, \label{pos}
\end{equation}
\begin{equation}
    {\left\| {{f\left(\mathbf{x}_{anc}\right)} - {f\left(\mathbf{x}_{neg}\right)}} \right\|_2^2} \leq \mathrm{t}_{neg},\label{neg}
\end{equation}
where $\mathrm{t}_{pos}$ and $\mathrm{t}_{neg}$ respectively denote the thresholds that determine whether a sample is hard positive or hard negative. Since the hardest samples are likely to be outliers which can degrade the training process of the triplet network, we empirically set the $\mathrm{t}_{pos}$ at 75 percentile of the L-2 distance and $\mathrm{t}_{neg}$ at 25 percentile of the L-2 distance. The final set of $\mathbf{x}_{pos}$ and $\mathbf{x}_{neg}$ samples is randomly selected from those samples threshold by~\eqref{pos} and~\eqref{neg}.

% Methods: ConvNets
\subsection{ConvNet structure}

The next step is to design a feature extractor which converts the input triplet data into feature vectors. In this work, we utilize the ConvNet structure~\cite{lecun1998gradient} as a feature extractor since the three-dimensional data format of our preprocessed input signal can be regarded as an image data format with multiple channels.

Our ConvNet structure (item (b) in Fig~\ref{fig1}) for the network consists of three convolutional layers and one fully-connected layer. The number of convolutional filters to be trained in each layer is empirically chosen as $\{64, 128, 256\}$, with fixed filter size of $3\times3$, each of stride 1. The Rectiﬁed Linear (ReLU) activation function and the Max-pooling layer are applied between each convolutional layer. Subsequently, the extracted features from the last convolutional layer are flattened into a vector before feeding into the fully-connected network.
The output vectors from the fully-connected layer are finally transformed using the sigmoid function followed by a L-2 normalization.

% Methods: Triplet loss
\subsection{The triplet loss}

The triplet loss function was first seen in \cite{hoffer2015deep} for training the triplet network.
\textbf{For selected anchor input sample $\mathbf{x}_{anc,i}$, the $i_{th}$ triplet input $\left\{\mathbf{x}_{anc,i}, \mathbf{x}_{pos,i}, \mathbf{x}_{neg,i}\right\}$ is generated by grouping positive input sample $\mathbf{x}_{pos,i}$ and the negative input sample $\mathbf{x}_{neg,i}$.
The generated triplet input is fed into ConvNet structure and make feature vectors $\left\{\mathbf{v}_{anc,i},\mathbf{v}_{pos,i},\mathbf{v}_{neg,i}\right\}$.}
The triplet loss function is formulated based on a summation of the difference between the positive distance (the L-2 distance between the anchor vector and the positive vector) and the negative distance (the L-2 distance between the anchor vector and the negative vector) as follows:
\begin{equation}
    loss = \sum_i^N max\left({ \left[ {\left\| {{\mathbf{v}_{anc,i}} - {\mathbf{v}_{pos,i}}} \right\|_2^2} - {\left\| {{\mathbf{v}_{anc,i}} - {\mathbf{v}_{neg,i}}} \right\|_2^2}  + \alpha \right]}, 0 \right),\label{triplet}
\end{equation}
where $N$ denotes the size of the mini-batch, ${\left\| . \right\|_2^2}$ denotes the L-2 distance and $\alpha$ denotes the preset margin.
The ConvNet structure using equation~\eqref{triplet} is trained to maximize the gap between the positive distance and the negative distance which should be larger than the margin $\alpha$.

@@

% Dataset
\subsection{Dataset}
 In order to evaluate the verification performance of the proposed system, the Wi-Fi CSI signature dataset \cite{moon2017air} with the user located at a single position is utilized in our experiments. The Wi-Fi CSI signature dataset consists of 2000 Wi-Fi CSI signature signals (4 directions $\times$ 10 samples $\times$ 50 identities) with sample size 500$\times$30$\times$6. We utilize only the absolute value from each complex CSI signal in our experiments.

@@

\subsubsection{Performance evaluation:}
The proposed system is evaluated under two cases: i) case I on comparison between the proposed system and other handcraft or deep learning-based methods based on the verification accuracy, and case II on detailed comparison between the proposed system and the deep learning-based methods using the receiver operating characteristic (ROC) curve and training loss curve. For Case I, existing handcraft methods such as the least squares error estimation (LSE), the principal components analysis (PCA) [22] with LSE, the support vector machine (SVM) with different kernel functions and a total error rate minimization which adopted the reduced multivariate polynomial model as basis function (TER-RM2) \cite{toh2003fingerprint,toh2008between}, the deep learning-based Siamese network \cite{koch2015siamese} and the baseline triplet network \cite{hoffer2015deep} are included for performance benchmarking. The Siamese network and the baseline triplet network utilized the same ConvNet structures with the proposed system and only differed in their input data style and loss function.

The verification performance of the proposed system and other methods are evaluated in terms of the Equal Error Rate (EER, \%) which are taken from averaging the results of five runs of two-fold cross-validation tests. Due to the memory constraint caused by the large data size, the deep learning-based methods utilized randomly sampled 9,500 negative pairs in the validation stage, which is the same number as the number of positive pairs.

\subsubsection{Network Structure and Parameter Settings:}
The multilayer feedforward network structure of KAR space learning is specified in Table~\ref{tab2}. With input data, size of 500$\times$30$\times$6, we set two network layers where the size of the each layer is 1024 and 16, respectively. Each layer is initialized with uniform distribution over [0, 1). We used $\sigma = {tan}^{-1}$ as an activation function following \cite{toh2018analytic}.

\begin{table}[]
    \caption{The network structure of KAR space learning.}\label{tab2}
    \centering
    \begin{tabular}{|l|l|l|}
    \hline
    Layer   & Size     & Activation \\ \hline
    Input   & 500$\times$30$\times$6 &            \\
    Fully-Connected 1 & 1$\times$1$\times$1024 & $\sigma = {tan}^{-1}$     \\
    Fully-Connected 2 & 1$\times$1$\times$16  & $\sigma = {tan}^{-1}$     \\
    Output  & 1$\times$1$\times$50   &            \\ \hline
    \end{tabular}
\end{table}

For the proposed system and the deep learning-based methods, we utilized the same ConvNet structure as specified in Table~\ref{tab1}. We trained the network starting with a learning rate of 0.00005 and a mini-batch size of 32. We optimized the loss by the Adam optimizer with L-2 penalty of 0.0002 except for the output layer. The output layer was regularized using an L-2 penalty of 0.0001. We initialized all network weights in the convolutional layers with normal distribution of zero-mean and standard deviation of 0.01. The biases were also initialized with a normal distribution of 0.5 mean and standard deviation 0.01. For the triplet networks, the hyper-parameter regulating triplet loss is empirically set at 0.1. The training epochs were set at 1,500 for all three deep learning-based algorithms. For the linear methods such as LSE,SVM and TER, the input signals were resized to 500 $\times$ 30 by averaging along the subcarrier axes due to limitation of hardware memory. For the PCA-LSE, the input dimension was reduced to 40.
 \begin{table}[]
    \caption{
        The structure of ConvNet model. For the convolution layer, kernel is specified as (m$\times$m) sized filter $\times$ \# of filters / \# of stride. For the max-pooling layer, (p$\times$p) sized pooling windows / \# of stride. The input sizes are denoted as rows $\times$ cols $\times$ \# of filters. 
    }\label{tab1}
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
    Layer     & Activation & Kernel / Stride & Input Size \\ \hline
    Conv 1    & ReLU       & (3$\times$3)$\times$64/1      & 500$\times$30$\times$6   \\
    MaxPool 1 &            & (2$\times$2)/1         & 500$\times$30$\times$64  \\
    Conv 2    & ReLU       & (3$\times$3)$\times$128/1     & 250$\times$15$\times$64 \\
    MaxPool 2 &            & (2$\times$2)/1         & 250$\times$15$\times$128 \\
    Conv 3    & ReLU       & (3$\times$3)$\times$256/1     & 125$\times$8$\times$128  \\
    MaxPool 3 &            & (2$\times$2)/1         & 125$\times$8$\times$256  \\
    Fully-Connected     & Sigmoid    & 16             & 63$\times$4$\times$256   \\
    L-2 Norm  &            &                 & 1$\times$1$\times$16    \\
    Concat    &            &                 & 1$\times$1$\times$16    \\ \hline
    \end{tabular}
\end{table}
